This chapter aims to give the reader enough of an overview of \Idris and dependent types to be able to understand the work done in the later chapters. A complete explanation of \Idris is beyond the scope of this chapter. This chapter borrows a lot of examples and explanations from the \Idris book (``Type-Driven Development with Idris'') \cite{brady_2017}.

\section{Types}

    \begin{code}[caption={In \Idris, the type of a function is specified using `\texttt{:}'}]
        anInt : Int
        anInt = 10
        
        aString : String
        aString = "foo"
        
        aBool : Bool
        aBool = False
    \end{code}

    Types classify values. In programming, we often come across types like \texttt{Int}, \texttt{String}, or \texttt{Bool}, which could be values like \texttt{10}, \texttt{"foo"}, or \texttt{False} respectively.

    \newpage

    \begin{code}[caption={Mismatching types}, escapeinside={(*}{*)}]
        aString : String
        aString = 10
        
        
        -- Compiler error --
        
          |
        2 | aString = 10
          |           ~~
        When checking right hand side of aString with expected type
        String
        
        String is not a numeric type
    \end{code}
    Types in \Idris are checked at compile time, meaning that if the types of the functions in a program do not match (for example, passing an \texttt{Int} where a \texttt{String} is required) then the program will not compile and the compiler will give a type-error.
    
    \begin{code}[label={des:no-cast}, caption={Values are not automatically cast}]
anInt : Int
anInt = 10

half : Double -> Double
half x = x / 2


> half anInt
(input):1:1-10:When checking an application of function Main.half:
        Type mismatch between
                Int (Type of anInt)
        and
                Double (Expected type)
    \end{code}
    
    \Idris is strongly typed, so the compiler will not automatically cast values or parameters. This means you cannot pass an \texttt{Int} to a function which requires a \texttt{Double} and have it automatically work, because \texttt{Int} is a different type, a different `category' from \texttt{Double}.
    
\section{Type Variables and Dependent Types}
    A different example of a type is a list of values. In languages like Java or Python, lists are parametrised over a type.
    \begin{code}[caption={The types of different lists in \Idris}]
        [1, 2, 3, 4, 5] : List Integer
        
        ["a", "b", "c"] : List String
        
        [True, False]   : List Bool
    \end{code}
    In \Idris this is still the case. You can have lists of \texttt{String}, lists of \texttt{Int}, or lists of \texttt{Bool} and these are different types. In general, you can have any \texttt{List elem} where \texttt{elem} is a \textit{type variable} representing the type of the elements of the list. \Idris provides an even more specific type of lists: \texttt{Vect}
    \begin{code}[caption={Example \texttt{Vect} types}]
        [1, 2, 3, 4, 5] : Vect 5 Int
        
        ["a", "b", "c"] : Vect 3 String
        
        [True, False]   : Vect 2 Bool
    \end{code}
    A \texttt{Vect}, short for ``vector'', is a list with a specific length. In general, you can have any \texttt{Vect n elem}. Here, \texttt{elem} is the same as in \texttt{List}, and \texttt{n} is the length of the list. Since the value of \texttt{n} \textit{depends} on the number of elements in the list, we refer to types like \texttt{Vect n elem} as \textit{dependent types} because its precise type depends on other values. Another example of a dependent type is \texttt{Matrix m n elem}, i.e. a matrix of \texttt{m} rows and \texttt{n} columns, with elements of type \texttt{elem}.
    \[
    \begin{pmatrix}
    1 & 2 \\
    3 & 4 \\
    5 & 6 \\
    \end{pmatrix}
    \hspace{50pt}
    \begin{pmatrix}
    ``1" & ``2" & ``3" \\
    ``4" & ``5" & ``6" \\
    \end{pmatrix}
    \]
    Here, the first matrix would have type \texttt{Matrix 3 2 Int} and the second matrix would have type \texttt{Matrix 2 3 String}.
    
    \begin{code}[caption={Types are \textit{first-class}}]
        ty : Type
        ty = Bool
        
        StringOrInt : Bool -> Type
        StringOrInt x = case x of
                            True => Int
                            False => String
    \end{code}
    
    In \Idris, types are \textit{first-class}. This means that they can be used without any restrictions: we can assign them to variables (\texttt{ty}), or have functions return them (\texttt{StringOrInt}). Types are used everywhere, in particular when defining functions.
    
\newpage
    

\section{Functions}
    In Listing \ref{des:no-cast} I defined a function \texttt{half}. Functions in \Idris consist of a \textit{type declaration} and a \textit{function definition}.
    \begin{code}[caption={The \texttt{half} function from Listing \ref{des:no-cast}}]
        half : Double -> Double
        half x = x / 2
    \end{code}
    The first line of the \texttt{half} function is its type declaration and the second line is the function definition. These can each be broken down into further parts. A function definition consists of:
    \begin{itemize}
        \item a function name, \texttt{half}
        \item a function type, \texttt{Double -> Double}
    \end{itemize}
    A function definition is a number of equations with:
    \begin{itemize}
        \item a right hand side, \texttt{half x}
        \item a left hand side, \texttt{x / 2}
    \end{itemize}
    \begin{code}[caption={A different definition of \texttt{half}}]
        half' : Double -> Double
        half' 0.0 = 0.0
        half' n = n / 2
    \end{code}
    In this different definition of \texttt{half}, \texttt{half'}, there are multiple functions definitions. A function can have varying definitions depending on its input. This is known as \textit{pattern matching} and is used extensively in \Idris. Here, it is not massively useful, as all it is doing is not bothering with the division, if the input is 0.

    \begin{code}[caption={A function with a function as an argument}]
        twice : (a -> a) -> a -> a
        twice f x = f (f x)
    \end{code}
    
    Functions in \Idris are first-class, so we can use them as arguments to other functions. The \texttt{twice} function's first argument is a function \texttt{f} which takes something of type \texttt{a} and produces something of type \texttt{a}. The second argument \texttt{x} is then something of type \texttt{a}. By not specifying a type and instead using a type variable, \texttt{twice} can be used on any type, as long as \texttt{x} has the same type as in the input to \texttt{f} and that \texttt{f} does not change the type of \texttt{x}.
    
    In addition to being first-class, functions in \Idris are \textit{pure} meaning that they do not have side effects (e.g. they do not do console I/O and they do not throw exceptions). This in turn means that for the same input, a function will always give the same output \cite{brady_2017}.


\section{Data Types}
    In addition to the given types, we can define our own data types. This is useful when we want to capture an aspect of the program we are writing which is likely to not be pre-defined in the \Idris prelude. Data types are defined using the \texttt{data} keyword. A data type is defined by a type constructor and one or more data constructors \cite{brady_2017}. Some of the included types are defined as data types, for example \texttt{List}.
    
    \begin{code}[caption={\texttt{List} as defined in the \Idris prelude}]
        data List : (elem : Type) -> Type where
            Nil  : List elem
            (::) : (x : elem) -> (xs : List elem) -> List elem 
    \end{code}
    A \texttt{List} of elements of type \texttt{Type} can be constructed by either constructing the empty list \texttt{Nil}, or an \texttt{x} of type \texttt{elem} (the type of all elements in the list) followed by a list of more elements of the same type.
    
    \begin{code}[caption={\texttt{Bool} as defined in the \Idris prelude}]
        data Bool = False
                  | True
    \end{code}
    Data types can be constructed with either equals and vertical bars (see the definition of \texttt{Bool} just above) or as functions, like in the definition of \texttt{List}. They are mostly interchangeable, but the \texttt{data ... -> Type where} syntax is more flexible and general, at the cost of also being more verbose.
    \\
    
    Data type definitions can also be recursive.


\section{Natural Numbers}\label{des:nats}
    \begin{code}[caption={Natural numbers as defined in the \Idris prelude}]
        data Nat = Z | S Nat
    \end{code}

    Natural numbers in \Idris are also a data record. They base case is the constant 0 (\texttt{Z} in Idris) and the other constructor is the \textit{successor function} \texttt{S}. From these two, all the natural numbers can be constructed:
    
    \begin{tabular}{r l}
        \texttt{Z} & $\mapsto 0$ \\ 
        \texttt{S Z} & $\mapsto 1$ \\ 
        \texttt{S (S Z)} & $\mapsto 2$ \\
        etc & \\
        \vdots & \\
    \end{tabular}
    \par
    The benefit of this is that it allows us to pattern match on numbers.
    
    \begin{code}[caption={Pattern matching on \texttt{Nat}s}]
        natFact : Nat -> Nat
        natFact Z = 1
        natFact (S k) = (S k) * fact k
    \end{code}
    By pattern matching on the possible constructors for \texttt{Nat}, we have captured the base case, and the recursive step for calculating the factorial of a number using the \texttt{natFact} function.
    Pattern matching also helps in terms of decidability.


\section{Decidable Equality}\label{idris:deceq}
    % TODO? split this into sections instead of sub-sections?
    \subsection{\texttt{Maybe} -- Handling uncertainty}
        Due to functions being pure, it can be difficult to reason about failure or uncertainty of a result being there.
        \begin{code}[caption={The \texttt{Maybe} data type}]
        data Maybe : (a : Type) -> Type where
            Nothing : Maybe a
            Just : (x : a) -> Maybe a
        \end{code}
        The \texttt{Maybe} data type captures this in the form of either having \texttt{Just} a value \texttt{x} or \texttt{Nothing}. We could use this to implement taking the n$^{th}$ element from a list.
        \begin{code}[caption={List indexing using \texttt{Maybe}}]
        getIndex : Nat -> List a -> Maybe a
        getIndex _ Nil = Nothing
        getIndex Z (x :: xs) = Just x
        getIndex (S k) (x :: xs) = getIndex k xs
        \end{code}
        Since we could either be given the empty list or run out of bounds, we use \texttt{Maybe} to reflect that the index may not be valid. If the index exists, we return \texttt{Just x} where \texttt{x} is the n$^{th}$ element, and if we run out of bounds we return \texttt{Nothing}.

    \subsection{The \texttt{(=)} data type}
        \begin{code}[label={des:refl-concept}, caption={The concept of an equality data type}]
            data (=) : a -> b -> Type where
                Refl : x = x
        \end{code}
        
        A very useful pre-defined data type is the \texttt{(=)} data type. It is not quite defined as a data type since `\texttt{=}' is a reserved symbol, instead it is part of the \Idris syntax. It is still a data type, just defined at a lower level of the language than the \texttt{data} construct. Its constructor \texttt{Refl} is short for `reflexive'. Reflexivity is a mathematical concept stating that every element in a set is related to itself. In logic, one would write $\forall x \in X : x R x$ where $X$ is a set of elements $x$ and $R$ is a binary relation. This further means that \texttt{=} also must be symmetric and transitive: if $x = y$ then $y = x$; and if $x = y$ and $y = z$ then $x = z$. In short, if two elements are reflexive, this means they are the same element. \textbf{[correct?]}
        
        \begin{code}[label={des:refl-egs}, caption={Examples of reflexivity}]
            Idris> the ("World" = "World") Refl
            Refl : "World" = "World"
            
            
            Idris> the (True = True) Refl
            Refl : True = True
            
            
            Idris> the (1 + 2 + 3 = 6) Refl
            Refl : 6 = 6
        \end{code}
        At the \Idris prompt, we can use the \texttt{the} function to create some example instances of \texttt{Refl}. The last example shows that \Idris evaluates the sides of a type before trying to construct the \texttt{Refl}.
        
        \newpage
        
        \begin{code}[label={des:not-refl}, caption={Things that are not reflexive}, escapeinside={(*}{*)}]
    Idris> the ("Foo" = "Bar") Refl
    (input):1:1-24:When checking argument value to function
    Prelude.Basics.the:
        Type mismatch between
            x = x (Type of Refl)
        and
            "Foo" = "Bar" (Expected type)
    
        Specifically:
            Type mismatch between
                "Foo"
            and
                "Bar"
    
    Idris> the (True = False) Refl
    (input):1:1-23:When checking argument value to function
    Prelude.Basics.the:
        Type mismatch between
            x = x (Type of Refl)
        and
            True = False (Expected type)
    
        Specifically:
            Type mismatch between
                True
            and
                False
    
    
    Idris> the (2 = 3) Refl
    (input):1:1-16:When checking argument value to function
    Prelude.Basics.the:
        Type mismatch between
            3 = 3 (Type of Refl)
        and
            2 = 3 (Expected type)
        
        Specifically:
            Type mismatch between
                3
            and
                2
        \end{code}
        If we try to instantiate \texttt{Refl} using elements which are not reflexive, the type-checker complains and tells us that it does not make sense and so we cannot create a \texttt{Refl} from things which are not reflexive. Combining the \texttt{=} data type with decidability proves to be very useful.

    \subsection{The \texttt{Void} data type}
    % TODO: rewrite `Void` here

    \subsection{Decidability}
        Decidability is more specific than \texttt{Maybe}. Instead of saying that we have just the value or nothing, \texttt{Dec} allows us to express that we can always decide whether a property holds or not for certain values.
        
        \begin{code}[caption={\texttt{Dec} as defined in the \Idris prelude}]
            data Dec : (prop : Type) -> Type where
                Yes : (prf : prop) -> Dec prop
                No  : (contra : prop -> Void) -> Dec prop
        \end{code}
        The definition of \texttt{Dec} seems similar to that of \texttt{Maybe}. \texttt{Yes (prf : prop) -> Dec prop} seems very similar to \texttt{Just : (x : a) -> Maybe a} and that is because they are. However, contrary to \texttt{Maybe}'s \texttt{Nothing} which simply was the absence of a value, \texttt{No} also holds a value: `\texttt{contra}'. This value is a proof that no value of the required type can exist. This is a much stronger statement as instead of saying ``the value is not there'' we have said ``the value can never be there''.
        % TODO: give examples of use-cases -- the book!

    % TODO: split this into each definition and explain
	\begin{code}[caption={Proving the equality between natural numbers \cite{brady_2017}},label={lst:decEq}]
    -- a proof that zero cannot be the successor of a natural number
    zeroNotSuc : (0 = S k) -> Void
    zeroNotSuc Refl impossible
    
    -- a proof that the successor of a natural number cannot be zero
    sucNotZero : (S k = 0) -> Void
    sucNotZero Refl impossible
    
    -- given a proof that two numbers ARE NOT equal and a proof that
    -- their successors ARE equal, return the empty type, i.e. a
    -- proof by contradiction
    noRec : (contra : (k = j) -> Void) -> (S k = S j) -> Void
    noRec contra Refl = contra Refl

    -- decidably describe whether two numbers are equal
    decEq : (a : Nat) -> (b : Nat) -> Dec (a = b)
    decEq Z Z = Yes Refl
    decEq Z (S j) = No zeroNotSuc
    decEq (S k) Z = No sucNotZero
    decEq (S k) (S j) = case decEq k j of
                            Yes prf => Yes (cong prf)
                            No contra => No (noRec contra)
	\end{code}

\section{Proofs}
    Proofs in \Idris rely on the type-checker and constructors. The reason \texttt{Refl} is a proof for equality is that it is defined as follows:
    \begin{code}[caption={Reflexivity as defined in the \Idris prelude \cite{brady_2017}}]
        data (=) : a -> b -> Type where
            Refl : x = x
    \end{code}
    The only constructor for \texttt{(=)} is \texttt{Refl} and \texttt{Refl} can only be constructed when the arguments are equal. Therefore, if a program using \texttt{Dec (t1 = t2)} type-checks, the only way that the program can return a \texttt{Yes Refl} is if the arguments to the function were truly equal. Similarly, the only way that the program could return a \texttt{No contra} is if there is no possible way that a \texttt{Refl} could be constructed; No possible way the arguments could be equal. This is a very common thing in proofs, and as such \Idris provides a number of things to help describe this.
    \\\par
    
    % TODO: Rewrite this - Alex didn't understand
    For counter-proofs, we have already seen the \texttt{Void} type (Section \ref{idris:dec}). \texttt{Void} is interesting because it is impossible to construct, meaning that given an instance of it we can do anything (e.g. prove \texttt{2 = 3}). The \texttt{void} function in \Idris captures this. It is defined as:
    \begin{code}
        void : Void -> a
    \end{code}
    Essentially, \texttt{Void} and \texttt{void} captures the absurdity of a contradiction. If we accept the contradiction to hold, we can then prove anything we want; We can construct any type we want. The \texttt{Uninhabited} interface allows us to specify instances which have no constructors and is defined as follows:
    \begin{code}[caption={The \texttt{Uninhabited} interface \cite{brady_2017}}]
        interface Uninhabited t where
            uninhabited : t -> Void
    \end{code}
    For example, we could define \texttt{Uninhabited} for \texttt{2 = 3}:
    \begin{code}
        implementation Uninhabited (2 = 3) where
            uninhabited Refl impossible
    \end{code}
    The \texttt{impossible} keyword specifies a case which would not type check and tells the \Idris type checker that it \textit{must} not type check. If the case specified as \texttt{impossible} could type-check, i.e. there \textit{is} a way to construct it, \Idris will print an error message and the program will not compile. Therefore, if a program using \texttt{impossible} compiles, we can be sure that it is truly not possible to construct something of that type.
    % TODO: `impossible` example
    This is in turn used to define the \texttt{absurd} function in the Prelude:
    \begin{code}[caption={The \texttt{absurd} function  \cite{brady_2017}}]
        absurd : Uninhabited t => (h : t) -> a
        absurd h = void (uninhabited h)
    \end{code}
    With \texttt{absurd} we have described exactly that: an absurdity. Hence, any constructors specified in an \texttt{Uninhabited} implementation allows a function to call \texttt{absurd}, thereby covering all cases and describing which of them cannot happen. Or rather, describing \textit{precisely which} can happen.
    
    % TODO: explain Uninhabited, absurd
